---
title: "AmpGram"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ranger)
library(kernlab)
library(biogram)
library(ggplot2)
library(hmeasure)
library(reshape2)

load("report_data.RData")
```

### Positive dataset

- Sequences from dbAMP: **12389**
- Selected those with experimental evidence: **4271**
- Removed those with non-standard amino acids (153): **4118** 
- Homology reduction with CD-HIT (90%): **2463**

### Negative dataset

- Sequences from UniProt (NOT antimicrobial NOT annotation:(type:transit) NOT antibacterial NOT antifungal NOT antiviral AND reviewed:yes)
- Combining all the input sequences into the one character vector
- Splitting the vector into parts of length equal to the length of the whole positive dataset 
- Each part is further splitted into smaller subsequences of lengths the same as lengths of sequences in the positive dataset
- To ensure the same number of sequences of the same length in both positive and negative datasets, cutted sequences are selected by choosing each subsequence (lengths defined by AMPs) from a randomly selected part of the larger vector (length of all AMPs

### Benchmark, training and testing dataset

- Sequences were divided  into **5 length groups** of equal sizes: [11:19], (19,26], (26,36], (36,60], (60,710]
- Benchmark dataset was created by selecting 1/10 of sequences from each length group: **494 sequences** (247 pos + 247 neg)
- The rest of sequences comprise train-test dataset **4432 sequences** (2216 pos + 2216 neg)

### Dataframe of mers

- All peptides are divided into 5 folds and cut into 10-mers: **87716 AMP mers** and **87599 non-AMP mers**

### Counting n-grams

- Generation and counting of n-grams 

### Cross validation

- 5-fold cross validation

### Prediction results

- Fraction of positive mers decreases with length of the peptide

```{r echo = FALSE, warning = FALSE}
ggplot(pred_len, aes(x = len, y = n_prop, fill = cfrac_true)) +
  geom_col() +
  facet_wrap(~ target, ncol = 1) +
  theme_bw() +
  scale_x_continuous(limits = c(0, 200))

```

```{r echo = FALSE, warning = FALSE}
filter(all_cvs, source_file == "(19,26]") %>% 
  group_by(source_peptide, target) %>% 
  summarise(fraction_true = mean(pred > 0.5),
            len = length(pred) + 9) %>% 
  filter(len <= 200) %>% 
  ggplot(aes(x = len, y = fraction_true, fill = target, color = target)) +
  stat_density2d(aes(alpha = ..level..), geom = "polygon", color = "black") +
  geom_smooth(method = "lm") +
  ylim(0, 1)
```

```{r echo = FALSE, warning = FALSE, fig.width = 10, fig.height = 10}
filter(all_cvs, source_file == "(19,26]" & group == "(36,60]") %>% 
  group_by(source_peptide, target) %>% 
  summarise(fraction_true = mean(pred > 0.5),
            len = length(pred) + 9) %>% 
  ggplot(aes(x = factor(len), y = fraction_true, fill = target, color = target)) +
  geom_violin(alpha = 0.5) +
  geom_point(aes(group = target), position = 'jitter') + 
  facet_wrap(~len, scales = "free_x")
```

20 randomly selected peptides from length group (36,60]

- Some non-AMP peptides possess fragments of sequences predicted as positive.

```{r fig.width = 10}

set.seed(2)
twenty_peps <- filter(all_cvs_pos, group == "(36,60]") %>% 
  pull(source_peptide) %>% 
  unique %>% 
  sample(20)

filter(all_cvs_pos, source_file == "(19,26]" & source_peptide %in% twenty_peps) %>% 
  ggplot(aes(x = mer_pos, y = pred,
             group = source_peptide)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.5, color = "red") +
  facet_grid(cfrac_true ~ target)

```

20 randomly selected peptides from length group (26,36]

- Some non-AMP peptides have very high fraction of positive mers.

```{r echo = FALSE}
set.seed(2)
twenty_peps <- filter(all_cvs_pos, group == "(26,36]") %>% 
  pull(source_peptide) %>% 
  unique %>% 
  sample(20)

filter(all_cvs_pos, source_file == "(19,26]" & source_peptide %in% twenty_peps) %>% 
  ggplot(aes(x = mer_pos, y = pred,
             group = source_peptide)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.5, color = "red") +
  facet_grid(cfrac_true ~ target)
```

### Statistics for the second layer of prediction

Following statistics are calculated for each peptide:

- fraction_true - fraction of positive mers,

- pred_mean - mean value of prediction,

- pred_median - median of prediction,

- n_peptide - number of mers in a peptide,

- n_pos - number of positive mers,

- pred_min - minimum value of prediction,

- pred_max - maximum value of prediction,

- longest_pos - the longest stretch of mers occuring consecutively and predicted as positive,

- n_pos_10 - number of streches comprising of at least 10 mers predicted as positive,

- frac_0_0.2 - fraction of mers with prediction in range [0,0.2],

- frac_0.2_0.4 - fraction of mers with prediction in range (0.2,0.4],

- frac_0.4_0.6 - fraction of mers with prediction in range (0.4,0.6],

- frac_0.6_0.8 - fraction of mers with prediction in range (0.6,0.8],

- frac_0.8_1 - fraction of mers with prediction in range (0.8,1]).


### Second layer - random forest

```{r echo = FALSE}

melt(perf_rf, id.vars = c(colnames(perf_rf)[c(1:2,4:10,13:25)]), measure.vars = c("AUC", "Sens", "Spec"), 
     variable.name = "Measure", value.name = "Value") %>% 
  ggplot(aes(x = source_file, y = Value, color = Measure)) +
  geom_point(position = "jitter") +
  facet_wrap(~ len_group)

```

### Second layer - other algorithms

```{r}

all_perf %>% 
  group_by(source_file, len_group, model) %>% 
  summarise(mean_AUC = mean(AUC)) %>% 
  ggplot(aes(x = source_file, y = mean_AUC, color = model)) +
  geom_point(position = "jitter") +
  facet_wrap(~ len_group)
```